---
title: ""
output:
  html_document:
    theme: yeti
    smart: true
    highlight: textmate
    toc: true

---


---
editor_options:
  markdown:
    wrap: 72
---

------------------------------------------------------------------------

------------------------------------------------------------------------

$$\\\\[0.2in]$$

## **GEOG0114 UKincome testsite**
$$\\\\[0.05in]$$
```{r}
# download packages
library(sf)
library(nngeo)
library(data.table)
library(tmap)
```

```{r}
uk_oas <- st_read("gdhi_data/nutsl3_gdhi_timeseries.shp", 
                      crs=27700)
uk_oas %>% 
  head()
```

```{r}
# plot(st_geometry(uk_oas))
# install.packages("rmapshaper")
library(rmapshaper)
library(tidyverse)
ms_uk_oas <- uk_oas %>% 
  ms_simplify(., keep = 0.05)
plot(ms_uk_oas$geometry)
```

```{r}
# plot(uk_oas$geometry)

```

### chosen OA test

```{r}
tm_shape(ms_uk_oas) +
  tm_borders(col="black") +
  tm_shape(ms_uk_oas[ms_uk_oas$objectid=="10",]) +
  tm_fill(col = "red")
```

#### Find the neighbors

```{r}
library(nngeo)
library(sf)
chosen_oa <- "10"
chosen_oa_neighbours <- st_nn(st_geometry(st_centroid(uk_oas[uk_oas$objectid==chosen_oa,])), 
                              st_geometry(st_centroid(uk_oas)),
                              sparse = TRUE,
                              k = 10,
                              maxdist = 600000) 
```

```{r}
class(chosen_oa_neighbours)
```

#### Get names of these neighbors

```{r}
neighbour_names <- uk_oas[chosen_oa_neighbours[[1]],]
neighbour_names <- neighbour_names$objectid
```

#### Neighbor map for NO.10

```{r}
tm_shape(ms_uk_oas) + 
  tm_borders() +
  # highlight only the neighbours
  tm_shape(ms_uk_oas[ms_uk_oas$objectid %in% neighbour_names,]) + 
  tm_fill(col = "green") +
  # highlight only the chosen OA
  tm_shape(ms_uk_oas[ms_uk_oas$objectid==chosen_oa,]) + 
  tm_fill(col = "red") +
  tm_shape(ms_uk_oas) + 
  # overlay the borders
  tm_borders(col="black")
```

#### queen function way

```{r}
st_queen <- function(a, b = a) st_relate(a, b, pattern = "F***T****")
chosen_oa_neighbours <- st_queen(st_geometry(ms_uk_oas[ms_uk_oas$objectid==chosen_oa,]),
                                 st_geometry(ms_uk_oas))

neighbour_names <- ms_uk_oas[chosen_oa_neighbours[[1]],]
neighbour_names <- neighbour_names$objectid

tm_shape(ms_uk_oas) + 
  tm_borders() +
  # highlight only the neighbours
  tm_shape(ms_uk_oas[ms_uk_oas$objectid %in% neighbour_names,]) + 
  tm_fill(col = "green") +
  tm_shape(ms_uk_oas[ms_uk_oas$objectid==chosen_oa,]) + 
  # highlight only the chosen OA
  tm_fill(col = "red") +
  tm_shape(ms_uk_oas) + 
  # overlay the borders
  tm_borders(col="black")

```

### processing income dataset

```{r}

tm_shape(ms_uk_oas)+
  tm_borders()+
  tm_fill("X2018", palette = "OrRd")+
  tm_layout("Income in UK in 2018", 
            bg.color = "white")+
  tm_legend(position = c("left", "bottom"))
```

here is the previous way of calculating neighbourhoodness for our chosen
OA but now we will add an extra argument that states we want a
non-sparse adjacency matrix returned

find only nearest neighbours to our chosen output area centroid, find
all within 500 meters select only the the centroid of our chosen output
area and all other areas (with st_centroid) we set the maximum number of
neighbours we want to find to "50" (with parameter k) we set the maximum
distance of calling an OA centroid a neigbour to "500" (with parameter
maxdist) we return a non-sparse matrix that tells us whether each OA is
a neighbour or not (with parameter sparse)

```{r}
chosen_oa_neighbours <- st_nn(st_geometry(st_centroid(uk_oas[uk_oas$objectid==chosen_oa,])), 
                              st_geometry(st_centroid(uk_oas)),
                              sparse = FALSE,
                              k = 10,
                              maxdist = 600000) 
```

### Global Moran's I

Some of the operations they will do will be similar to the examples
shown earlier, but the way they assign and store variables makes it much
quicker to run complex spatial operations.

```{r}
# for doing spatial operations (now mostly superseded by sf)
library(sp)

# for our neighbourhood and Moran's I analysis
# install.packages("deldir")
library(spdep)
```

#### Find neighbours

```{r}
# neighbors
uk_oas_sp <- as_Spatial(uk_oas, IDs=uk_oas$objectid)
uk_oas_nb <- poly2nb(uk_oas_sp, row.names=uk_oas_sp$objectid)

```
```{r}
summary(uk_oas_nb)
```


```{r}
str(uk_oas_nb,list.len=10)
```

#### List weights object

```{r}
# create the list weights object
nb_weights_list <- nb2listw(uk_oas_nb, style='B', zero.policy = TRUE)

# have a look at the class
class(nb_weights_list)
```

```{r}
# so it's nb AND weights

# now use that to create a quick Moran's I
moran(uk_oas_sp$X2018, 
      nb_weights_list, 
      n=length(nb_weights_list$neighbours), 
      S0=Szero(nb_weights_list),
      zero.policy = TRUE)
```

#### pseudo p-value

```{r}
# run it 599 times
mc_model <- moran.mc(uk_oas_sp$X2018, nb_weights_list, zero.policy = TRUE, nsim=599)

# what do we get?
mc_model
```

### Local Moran's I

#### Find neighbours

```{r}
# you need the nb object and the nb and weights list
# make the nb
# uk_oas_nb <- poly2nb(uk_oas_sp, row.names=uk_oas_sp$objectid)

```

#### List weights object

```{r}
# create the list weights object
# but importantly with the row stadardisation this time
nb_weights_list <- nb2listw(uk_oas_nb, zero.policy = TRUE, style='W')

# use the localmoran() function
local_moran_uk_oa_income <- localmoran(uk_oas_sp$X2018, zero.policy = TRUE, nb_weights_list)
```

#### Rescale that variable

```{r}
# rescale that variable!
uk_oas_sp$scale_n_income <- scale(uk_oas_sp$X2018)
```

#### Spatial lag variable

```{r}
# create a spatial lag variable and save it to a new column
uk_oas_sp$lag_scale_n_income <- lag.listw(nb_weights_list, uk_oas_sp$scale_n_income, zero.policy = TRUE)
```

#### Convert to sf

```{r}
# convert to sf
uk_oas_moran_stats <- st_as_sf(uk_oas_sp)
```

#### Without statistical significance version

```{r}
uk_oas_moran_stats$quad_non_sig <- ifelse(uk_oas_moran_stats$scale_n_income > 0 & 
                                            uk_oas_moran_stats$lag_scale_n_income > 0, 
                                          "high-high", 
                                          ifelse(uk_oas_moran_stats$scale_n_income <= 0 & 
                                                   uk_oas_moran_stats$lag_scale_n_income <= 0,
                                                 "low-low", 
                                                 ifelse(uk_oas_moran_stats$scale_n_income > 0 & 
                                                          uk_oas_moran_stats$lag_scale_n_income <= 0, 
                                                        "high-low", 
                                                        ifelse(uk_oas_moran_stats$scale_n_income <= 0 & 
                                                                 uk_oas_moran_stats$lag_scale_n_income > 0,
                                                               "low-high",NA))))
```

```{r}
library(ggplot2)
```

```{r}
# plot the results without the satistical significance
ggplot(uk_oas_moran_stats, aes(x = scale_n_income,
                               y = lag_scale_n_income,
                               color = quad_non_sig)) +
  geom_vline(xintercept = 0) + # plot vertical line
  geom_hline(yintercept = 0) + # plot horizontal line
  xlab("Scaled income (n)") +
  ylab("Lagged Scaled income (n)") +
  labs(colour="Relative to neighbours") +
  geom_point()
```

```{r}
# map all of the results here
tm_shape(uk_oas_moran_stats) +
  tm_fill(col = "quad_non_sig")
```

#### Statistical significance version

```{r}
# run all of these through to assign variables

# for the statistical significance version assign a level 
# of statistical significance for the p value, column 5 of the local moran model
sig_level <- 0.1

# version with significance value
uk_oas_moran_stats$quad_sig <- ifelse(uk_oas_moran_stats$scale_n_income > 0 & 
                                        uk_oas_moran_stats$lag_scale_n_income > 0 & 
                                        local_moran_uk_oa_income[,5] <= sig_level, 
                                      "high-high", 
                                      ifelse(uk_oas_moran_stats$scale_n_income <= 0 &
                                               uk_oas_moran_stats$lag_scale_n_income <= 0 & 
                                               local_moran_uk_oa_income[,5] <= sig_level, 
                                             "low-low", 
                                             ifelse(uk_oas_moran_stats$scale_n_income > 0 & 
                                                      uk_oas_moran_stats$lag_scale_n_income <= 0 & 
                                                      local_moran_uk_oa_income[,5] <= sig_level, 
                                                    "high-low", 
                                                    ifelse(uk_oas_moran_stats$scale_n_income <= 0 & 
                                                             uk_oas_moran_stats$lag_scale_n_income > 0 & 
                                                             local_moran_uk_oa_income[,5] <= sig_level,
                                                           "low-high",
                                                           ifelse(local_moran_uk_oa_income[,5] > sig_level,
                                                                  "not-significant",
                                                                  "not-significant")))))
```

```{r}
# plot the results nnw with the satistical significance
ggplot(uk_oas_moran_stats, aes(x = scale_n_income, 
                                   y = lag_scale_n_income, 
                                   color = quad_sig)) +
  geom_vline(xintercept = 0) + # plot vertical line
  geom_hline(yintercept = 0) + # plot horizontal line
  xlab("Scaled income (n)") +
  ylab("Lagged Scaled income (n)") +
  labs(colour="Relative to neighbours") +
  geom_point()
```

```{r}
# map only the statistically significant results here
tm_shape(uk_oas_moran_stats) +
  tm_fill(col = "quad_sig")
```

